maxent model avec xvalidation (year-based test training)

### 2019-05-13 src/maxent_avec_xvalidation.Rmd
probability categorised maps and plot finished

#```{r train and test inc AIC for reg val}
presyrs <-  #this is to get years with complete presence data

@@

#```

@@

>>>>

```{r train and test inc AIC for reg val}
presyrs <-  #this is to get years with complete presence data

@@

@@

#probability selected maps

```{r habitat classification on probability - 3 classifications}

binprob <- function(x){
  ifelse(x >= 0.8, 3,
  ifelse(x >= 0.7 & x <0.8, 2,
  ifelse(x >= 0.6 & x <0.7, 1,
  ifelse(x < 0.6, 0, NA)))) 
}

  modelmonth <- cbind(paste0(mthna, unique(prback$year)))
  clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)

  for (modtime in 1:nrow(modelmonth)){
    timeslice <- modelmonth[modtime]
    yr <- str_sub(timeslice, -4)
    for (cum in 1:length(clogasc_list)){
      cumascyr <- str_sub(clogasc_list[cum], start = 44, end = 47) #selects the year from the file name
      if(yr == cumascyr){
        cumasc <- raster(clogasc_list[cum])
        ascname <- names(cumasc)
        probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
        writeRaster(probbin, paste0("../output/maxent/predictions/", mthna,  "/binary_prob_", ascname, ".asc"), overwrite = TRUE)
        plot(probbin, png(paste0("../output/maxent/predictions/", mthna, "/binary_prob_", ascname, ".png")), overwrite = TRUE)
        dev.off() # stops automatic saving of the plot to a png
      }
    }
  }
```

```{r habitat classification on probability - high only}

binprob <- function(x){
  ifelse(x >= 0.8, 1,
  ifelse(x < 0.8, 0, NA)) 
}

modelmonth <- cbind(paste0(mthna, unique(prback$year)))
clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)

  for (modtime in 1:nrow(modelmonth)){
    timeslice <- modelmonth[modtime]
    yr <- str_sub(timeslice, -4)
    for (cum in 1:length(clogasc_list)){
      cumascyr <- str_sub(clogasc_list[cum], start = 44, end = 47) #selects the year from the file name
      if(yr == cumascyr){
        cumasc <- raster(clogasc_list[cum])
        ascname <- names(cumasc)
        probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
        writeRaster(probbin, paste0("../output/maxent/predictions/", mthna,  "/binary_highprob_", ascname, ".asc"), overwrite = TRUE)
        plot(probbin, png(paste0("../output/maxent/predictions/", mthna, "/binary_highprob_", ascname, ".png")), overwrite = TRUE)
        dev.off() # stops automatic saving of the plot to a png
      }
    }
  }

```{r habitat classification on probability - med only}
binprob <- function(x){
  ifelse(x >= 0.7 & x <0.8, 1,
  ifelse(x < 0.7, 0, NA)) 
}

modelmonth <- cbind(paste0(mthna, unique(prback$year)))
clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)

  for (modtime in 1:nrow(modelmonth)){
    timeslice <- modelmonth[modtime]
    yr <- str_sub(timeslice, -4)
    for (cum in 1:length(clogasc_list)){
      cumascyr <- str_sub(clogasc_list[cum], start = 44, end = 47) #selects the year from the file name
      if(yr == cumascyr){
        cumasc <- raster(clogasc_list[cum])
        ascname <- names(cumasc)
        probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
        writeRaster(probbin, paste0("../output/maxent/predictions/", mthna,  "/binary_medprob_", ascname, ".asc"), overwrite = TRUE)
        plot(probbin, png(paste0("../output/maxent/predictions/", mthna, "/binary_medprob_", ascname, ".png")), overwrite = TRUE)
        dev.off() # stops automatic saving of the plot to a png
      }
    }
  }
```

```{r habitat classification on probability - low only}
binprob <- function(x){
  ifelse(x >= 0.6 & x <0.7, 1,
  ifelse(x < 0.6, 0, NA)) 
}

modelmonth <- cbind(paste0(mthna, unique(prback$year)))
clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)

  for (modtime in 1:nrow(modelmonth)){
    timeslice <- modelmonth[modtime]
    yr <- str_sub(timeslice, -4)
    for (cum in 1:length(clogasc_list)){
      cumascyr <- str_sub(clogasc_list[cum], start = 44, end = 47) #selects the year from the file name
      if(yr == cumascyr){
        cumasc <- raster(clogasc_list[cum])
        ascname <- names(cumasc)
        probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
        writeRaster(probbin, paste0("../output/maxent/predictions/", mthna,  "/binary_lowprob_", ascname, ".asc"), overwrite = TRUE)
        plot(probbin, png(paste0("../output/maxent/predictions/", mthna, "/binary_lowprob_", ascname, ".png")), overwrite = TRUE)
        dev.off() # stops automatic saving of the plot to a png
      }
    }
  }
```

now you need to do summary maps....

```{r high probability map per year 2D representation}
binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_highprob_cloglog.*.asc", full.names = TRUE)
yrlayer <- data.frame(matrix(ncol = 2, nrow = length(binaryprob_list)))
colnames(yrlayer) <- c("layer", "year")

for (yrlyr in 1:length(binaryprob_list)){
  yearoflyr <- str_sub(binaryprob_list[yrlyr], start = 60, end = 63) #selects the year from the file name
  yrlayer$year[yrlyr] <- yearoflyr
  yrlayer$layer[yrlyr] <- binaryprob_list[yrlyr]
}

for (binyr in (1:length(y_list))){
  binstack <- stack(y_list[[binyr]]$layer)
  year <- y_list[[binyr]]$year[1]
  stacksum <- sum(binstack, na.rm = TRUE)
  writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/highprob_2d_", year, ".asc"), overwrite=TRUE)
  plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/highprob_2d_", year, ".png")))
  dev.off()
}

```{r med probability map per year 2D representation}
binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_medprob_cloglog.*.asc", full.names = TRUE)
yrlayer <- data.frame(matrix(ncol = 2, nrow = length(binaryprob_list)))
colnames(yrlayer) <- c("layer", "year")

for (yrlyr in 1:length(binaryprob_list)){
  yearoflyr <- str_sub(binaryprob_list[yrlyr], start = 59, end = 62) #selects the year from the file name
  yrlayer$year[yrlyr] <- yearoflyr
  yrlayer$layer[yrlyr] <- binaryprob_list[yrlyr]
}

for (binyr in (1:length(y_list))){
  binstack <- stack(y_list[[binyr]]$layer)
  year <- y_list[[binyr]]$year[1]
  stacksum <- sum(binstack, na.rm = TRUE)
  writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/medprob_2d_", year, ".asc"), overwrite=TRUE)
  plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/medprob_2d_", year, ".png")))
  dev.off()
}

```{r low probability map per year 2D representation}
binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_lowprob_cloglog.*.asc", full.names = TRUE)
yrlayer <- data.frame(matrix(ncol = 2, nrow = length(binaryprob_list)))
colnames(yrlayer) <- c("layer", "year")

for (yrlyr in 1:length(binaryprob_list)){
  yearoflyr <- str_sub(binaryprob_list[yrlyr], start = 59, end = 62) #selects the year from the file name
  yrlayer$year[yrlyr] <- yearoflyr
  yrlayer$layer[yrlyr] <- binaryprob_list[yrlyr]
}

for (binyr in (1:length(y_list))){
  binstack <- stack(y_list[[binyr]]$layer)
  year <- y_list[[binyr]]$year[1]
  stacksum <- sum(binstack, na.rm = TRUE)
  writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/lowprob_2d_", year, ".asc"), overwrite=TRUE)
  plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/lowprob_2d_", year, ".png")))
  dev.off()
}

now average probability split maps - again low, med, high and all

```{r habitat classification on average probability - 3 classifications}

binprob <- function(x){
  ifelse(x >= 0.8, 3,
  ifelse(x >= 0.7 & x <0.8, 2,
  ifelse(x >= 0.6 & x <0.7, 1,
  ifelse(x < 0.6, 0, NA)))) 
}

clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^avg_cloglog_.*.asc", full.names = TRUE)

    for (cum in 1:length(clogasc_list)){
      cumasc <- raster(clogasc_list[cum])
      ascname <- names(cumasc)
      probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
      writeRaster(probbin, paste0("../output/maxent/predictions/", mthna,  "/binary_avgprob_", ascname, ".asc"), overwrite = TRUE)
      plot(probbin, png(paste0("../output/maxent/predictions/", mthna, "/binary_avgprob_", ascname, ".png")), overwrite = TRUE)
      dev.off() # stops automatic saving of the plot to a png
    }

```{r habitat classification on average probability - high}

binprob <- function(x){
  ifelse(x >= 0.8, 1,
  ifelse(x < 0.8, 0, NA))  
}

clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^avg_cloglog_.*.asc", full.names = TRUE)

    for (cum in 1:length(clogasc_list)){
      cumasc <- raster(clogasc_list[cum])
      ascname <- names(cumasc)
      probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
      writeRaster(probbin, paste0("../output/maxent/predictions/", mthna,  "/binary_avg_highprob_", ascname, ".asc"), overwrite = TRUE)
      plot(probbin, png(paste0("../output/maxent/predictions/", mthna, "/binary_avg_highprob_", ascname, ".png")), overwrite = TRUE)
      dev.off() # stops automatic saving of the plot to a png
    }

```{r habitat classification on average probability - med}

binprob <- function(x){
  ifelse(x >= 0.7 & x <0.8, 1,
  ifelse(x < 0.7, 0, NA)) 
}

clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^avg_cloglog_.*.asc", full.names = TRUE)

    for (cum in 1:length(clogasc_list)){
      cumasc <- raster(clogasc_list[cum])
      ascname <- names(cumasc)
      probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
      writeRaster(probbin, paste0("../output/maxent/predictions/", mthna,  "/binary_avg_medprob_", ascname, ".asc"), overwrite = TRUE)
      plot(probbin, png(paste0("../output/maxent/predictions/", mthna, "/binary_avg_medprob_", ascname, ".png")), overwrite = TRUE)
      dev.off() # stops automatic saving of the plot to a png
    }

```{r habitat classification on average probability - low}

binprob <- function(x){
  ifelse(x >= 0.6 & x <0.7, 1,
  ifelse(x < 0.6, 0, NA)) 
}

clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^avg_cloglog_.*.asc", full.names = TRUE)

    for (cum in 1:length(clogasc_list)){
      cumasc <- raster(clogasc_list[cum])
      ascname <- names(cumasc)
      probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
      writeRaster(probbin, paste0("../output/maxent/predictions/", mthna,  "/binary_avg_lowprob_", ascname, ".asc"), overwrite = TRUE)
      plot(probbin, png(paste0("../output/maxent/predictions/", mthna, "/binary_avg_lowprob_", ascname, ".png")), overwrite = TRUE)
      dev.off() # stops automatic saving of the plot to a png
    }

```{r high probability average map per year 2D representation}
binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_highprob.*.asc", full.names = TRUE)

binstack <- stack(binaryprob_list)
stacksum <- sum(binstack, na.rm = TRUE)
writeRaster(stastaccksum, paste0("../output/maxent/predictions/", mthna, "/binary_highavg_2d.asc"), overwrite=TRUE)
plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/binary_highprob_avg_2d.png")))
dev.off()
```
```{r med probability average map per year 2D representation}
binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_medprob.*.asc", full.names = TRUE)

binstack <- stack(binaryprob_list)
stacksum <- sum(binstack, na.rm = TRUE)
writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/binary_medavg_2d.asc"), overwrite=TRUE)
plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/binary_medprob_avg_2d.png")))
dev.off()
```

```{r low probability average map per year 2D representation}
binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_lowprob.*.asc", full.names = TRUE)

binstack <- stack(binaryprob_list)
stacksum <- sum(binstack, na.rm = TRUE)
writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/binary_lowavg_2d.asc"), overwrite=TRUE)
plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/binary_lowprob_avg_2d.png")))
dev.off()
```

```{r calc no presence cells per layer (avg probability)} 
ascdepth_list_high <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_highprob_.*asc", full.names = TRUE)
ascdepth_list_med <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_medprob_.*asc", full.names = TRUE)
ascdepth_list_low <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_lowprob_.*asc", full.names = TRUE)

binaryavgstack <- stack(ascdepth_list_high)
stackstathigh <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
stackstathigh[ ,1] <- c(1:49)
colnames(stackstathigh) <- c("layer", "high_cell_sum")

for (lyr in 1:nlayers(binaryavgstack)){
  lyrsum <- cellStats(binaryavgstack[[lyr]], sum)
  stackstathigh$high_cell_sum[lyr] <- lyrsum
}

binaryavgstack <- stack(ascdepth_list_med)
stackstatmed <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
stackstatmed[ ,1] <- c(1:49)
colnames(stackstatmed) <- c("layer", "med_cell_sum")

for (lyr in 1:nlayers(binaryavgstack)){
  lyrsum <- cellStats(binaryavgstack[[lyr]], sum)
  stackstatmed$med_cell_sum[lyr] <- lyrsum
}

binaryavgstack <- stack(ascdepth_list_low)
stackstatlow <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
stackstatlow[ ,1] <- c(1:49)
colnames(stackstatlow) <- c("layer", "low_cell_sum")

for (lyr in 1:nlayers(binaryavgstack)){
  lyrsum <- cellStats(binaryavgstack[[lyr]], sum)
  stackstatlow$low_cell_sum[lyr] <- lyrsum
}

stackstat <- cbind(stackstathigh, stackstatmed[ , 2], stackstatlow[ ,2])
colnames(stackstat) <- c("layer", "high", "med", "low")
stackstat <- melt(stackstat, id = "layer")

write.csv(stackstat, paste0("../output/maxent/predictions/", mthna, "/probability_cellcount_bydepth_avg.csv"), row.names = FALSE)
```

```{r graph by depth - prob plot (avg)}
bydepth <- ggplot(data = stackstat, aes(x = layer, y = value, fill = variable)) + geom_bar(stat = "identity") +   scale_x_reverse() +
coord_flip()
png(paste0("../output/maxent/predictions/", mthna, "/depthplot_probavg.png"))
print(bydepth)
dev.off()
```
